---
title: "class15"
author: "Yushan LIU"
date: "February 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#import countData and colData into R

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
dim(head)
head(counts)
```
```{r}
head(metadata)
```
```{r}
colnames(counts)
```
```{r}
metadata$id
```

Check for corespondence of metadata and countdata
```{r}
all(colnames(counts)[-1] == metadata$id)
```


First task find which columns in **counts** correspond to the control conditions (i.e. no drug!). Then we can calculate the mean count value for each gene.

```{r}
control.inds <-metadata$dex == "control"
control.ids <-metadata$id[control.inds]
head(counts[,control.inds-1])
```

Calculate the mean counts value for all genes
```{r}
control.mean<-rowSums(counts[control.ids])/ncol(counts[,control.ids])
names(control.mean) <- counts$ensgene
head(control.mean)
```

```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums(counts[,treated$id])/ncol(counts[,treated$id])

```

Combine the mean count values for control and treated into a new data.frame called meancounts!

```{r}
meancounts <- data.frame(control.mean,treated.mean)
```

```{r}
colSums(meancounts)
```

Now we can plot the drug treated vs non-drug control count values

```{r}
plot(meancounts$control.mean,meancounts$treated.mean)
```

```{r}
plot(log="xy",meancounts$control.mean,meancounts$treated.mean)
```

We can find candidate differentially expressed genes by looking for genes with a large change between control and dex-treated samples. We usually look at the log2 of the fold change, because this has better mathematical properties.
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
```{r}
toy <- c(1,2,0,4,0)
toy == 0
which(toy==0)
```

```{r}
toy <- meancounts[1:6,]
toy[,1:2] == 0
unique(which(toy[,1:2]==0,arr.ind=TRUE)[,1])
```

find zero 
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind)
```

```{r}
sum(down.ind)
```


There are `sum(up.ind)` up related genes and `sum(down.ind)` down regulated genes.


```{r}
mycounts[up.ind,]
```


Adding data

```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```

Use merge function to add data

```{r}
mycounts.annotated<-merge(x=mycounts,y=anno,by.x="row.names",by.y="ensgene")
head(mycounts.annotated)
```

```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```


```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```
```{r}
summary(res)
```

```{r}
res01 <- results(dds, alpha=0.01)
summary(res01)
```

```{r}
mycols <-rep("gray",nrow(res01))
mycols[abs(res01$log2FoldChange) >2]<-"blue"
mycols[]
plot(res01$log2FoldChange,-log(res01$padj),col=mycols)
```

