---
title: "class09"
author: "Yushan LIU"
date: "February 7, 2019"
output: html_document
---

Preparing the data
```{r}
# Save your input data file to a new 'data' directory
fna.data <- "WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data)
wisc.df
```
How many samples (i.e. patients) are in this data-set



```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])
```

```{r}
# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id
head(wisc.data)
```

Finally, setup a separate new vector called diagnosis to be 1 if a diagnosis is malignant ("M") and 0 otherwise. Note that R coerces TRUE to 1 and FALSE to 0.
```{r}
# Create diagnosis vector by completing the missing code
#tmp<-rep(0,nrow(wisc.df))
diagnosis <- as.numeric(wisc.df$diagnosis=='M')
diagnosis
```

Q1. How many observations are in this dataset?
Q2. How many variables/features in the data are suffixed with _mean?
Q3. How many of the observations have a malignant diagnosis?

```{r}
#1
nrow(wisc.data)
#2
inds<-grep("_mean",colnames(wisc.data))
colnames(wisc.data)[inds]
length(inds)
#3
sum(diagnosis)
```

```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
```

##Lets do PCA

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data,scale=TRUE)
```
```{r}
# Look at summary of results
summary(wisc.pr)
```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
0.4427
Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
3
Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
7

##Interpreting PCA results

```{r}
biplot(wisc.pr)
```
```{r}
# Scatter plot observations by components 1 and 2
plot( wisc.pr$x[,1], wisc.pr$x[,2],col=diagnosis+1,
     xlab = "PC1", ylab = "PC2")
```
### Variance captured in each PC

This info is in the $sdev component of our PCA result
```{r}
varance <- wisc.pr$sdev^2
pve<-round((varance/sum(varance))*100,1)
```
```{r}
plot(pve,type="o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )

```


##Hierarchical clustering of case data

For hiearchial clustering 
```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```

```{r}
data.dist <- dist(data.scaled)
wisc.hclust <-  hclust(data.dist，method="complete")
```

```{r}
plot(wisc.hclust)
```

### Cluster in PCA space
For clustering we need?
1. Distance matrix
2. Clustering function
3. cutree
```{r}
plot(wisc.pr$x[,1],wisc.pr$x[,2],col=diagnosis+1)
```

```{r}
pc.dist<-dist(wisc.pr$x[,1:2])
pc.hclust<-hclust(pc.dist,method="ward.D2")
plot(pc.hclust)
```

```{r}
grps3<-cutree(pc.hclust,k=3)
table(grps3)
```
```{r}
table(grps3,
      diagnosis)
```

##Prediction
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```
```{r}
plot(wisc.pr$x[,1:2], col=grps3)
points(npc[,1], npc[,2], col="blue", pch=16)
```


