---
title: "class13:Structure based drug discovery"
author: "Yushan LIU"
date: "February 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Prepare HIV-protease molecule for docking

First get a PDB structure from the online database then clean it up for docking
```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```
Read file into R and clean it up (remove non-protein bits) and also we will save the ligand into separate file
```{r}
hiv <- read.pdb(file.name)
hiv
```

Q1: What is the name of the two non protein resid values in this structure?  What does resid correspond to and how would you get a listing of all reside values in this structure? 
HOH (127), MK1 (1)
```{r}
prot <- trim.pdb(hiv,"protein")
lig  <- trim.pdb(hiv, "ligand")
```
```{r}
write.pdb(prot,file="1hsg_protein.pdb")
write.pdb(lig,file="1hsg_lig.pdb")
```

```{r}
library(bio3d) 
res <- read.pdb("all.pdbqt", multi=TRUE) 
write.pdb(res, "results.pdb") 
```

```{r}
# 
res <- read.pdb("all.pdbqt", multi=TRUE) 
ori <- read.pdb("1hsg_lig.pdbqt") 
rmsd(ori, res)
```

Q4:  Qualitatively, how good are the docks?  Is the crystal binding mode reproduced? Is it the best conformation according to AutoDock Vina?

Q5:  Quantitatively how good are the docks?  Is the crystal binding mode reproduced within 1Å RMSD for all atoms? 

Q6:  How would you determine the RMSD for heavy atoms only (i.e. non hydrogen atoms)? HINT: The atom.select() function will be of help here along with the selection string “noh” for no hydrogens.

```{r}
pdb <- read.pdb("1HEL")
m<-nma(pdb)
plot(m)
```
```{r}
# Visualize NMA results 
mktrj(m, mode=7, file="nma_7.pdb")

```


